Online Social Cohesion Reflects Real-World Group Action in Syria during the Arab Spring

### 2020-06-01 manuscript.Rmd
More edits to the Cohesion section

### Lexical Cohesion: Connectedness of Words between Interlocutors

@@

## Social and Physical Cohesion
In turn, other studies show that linguistic alignment leads to increased social
connectedness and solidarity [@Miles_Nind_Macrae_2009], which is associated with
increased motivation for participation [@Walton_Cohen_Cwir_Spencer_2012],
facilitating successful coordination of group actions toward shared goals
[@Fusaroli_Bahrami_Olsen_Roepstorff_Rees_Frith_Tylén_2012;
@Richardson_Dale_2005]. This research emphasizes both the importance of and
connections between linguistic alignment and behavioral alignment. The more
people interact, the more similar their language and linguistic patterns become.
This holds true for face-to-face interactions as well as online communication
[@Riordan_Dale_Kreuz_Olney_2011}. Given this, we should expect to see increased
linguistic cohesion amongst people participating in online discussions about an
idea, event, or individual. Given that social media reports in real time about
actual events taking place, i.e., protests or demonstrations in which people are
assembled together in a location, or attacks that require coordination between
members of a unit, the cohesion in social media should be expected to correlate
with the social cohesion of the group [@Zeitzoff_2011}.

A socially cohesive group is inherently more threatening to the existing regime
because it reflects the professionalization of potential alternate leadership,
and collective opposition momentum. Groups with high social cohesion are more
difficult to divide and conquer, and more cohesive messages may also garner
increased international support. A paper published in Nature argues that global
insurgency movements all exhibit a similar pattern to their social structure
[@Bohorquez_Gourley_Dixon_Spagat_Johnson_2009}. The authors assert that each
violent group has a stable state, and that exogenous events can create
perturbations that push the group to become more cohesive, or alternatively to
become more fragmented. Rather than frenetic, patchwork messages via Twitter,
cohesive messages indicate increased organizational structure, which should
incentivize a response from the government. What process is the driving force,
though? Does an increase in linguistic cohesion precede or follow increased
repression by the government? Do movements get stronger first, and then the
government responds, or does the government increase its offensive measures, and
thus social groups have a clear opponent around whom they can coalesce both
socially and linguistically?

Thus, verbal synchrony leads to physical synchrony. Shared language in an online
environment can influence individuals’ propensity to engage in similar
coordinated behavior. In the case of the Arab Spring, this may mean social
mobilization in public spaces to protest against government brutality. Words in
a passage of text – or series of tweets – have some degree of semantic relation.
Individuals engaged in a conversation, whether in real life or online, have the
tendency to use similar language as they interact, increasing the cohesion of
their language. As an extension, multiple individuals collectively communicating
can also demonstrate linguistic alignment. In the context of social conflict,
the linguistic cohesion of participants in the online conversation should
correspond to the degree of social cohesion on the ground. We link these
concepts of social and linguistic cohesion together to demonstrate how language
serves as an indicator of the degree of social mobilization, using Twitter feeds
during the spring of 2012 in Syria.

@@

Physical and social cohesion in the world can be modeled using event data which
encodes who does what to whom on a given day. This could include events such as
the government (GOV) imposing a blockade or fighting with artillery or tanks on
a civilian group (CIV). CAMEO (Conflict and Mediation Event Observations) encode
a set of activities with numerical values, rescaled to reflect a
conflict-cooperation scale [@Schrodt_Hall_2006; @Goldstein_1992]. These event
codes include a range of behaviors, from making public statements and appeals,
to demanding, threatening, protesting, and using unconventional mass violence.
Steinert-Threlkeld (2017) modeled a similar relationship between social media
and social mobilization using a Gini coefficient measure, rather than a tf-idf
(term frequency-inverse document frequency) measure alongside event data
[@Ramos_2003; @Threlkeld_2017]. This pathbreaking work demonstrates that
participants on the fringe of a social movement contribute substantially to the
overall level of protest. We add to this discussion by demonstrating how
sociolinguistic cohesion evolves in response to negative and positive exogenous
events.

### Cohesion in Social Movements and Civil Wars
People join social and political resistance movements because they feel a sense
of connection to the cause, and because they have been sufficiently convinced to
participate. Given that people are generally risk averse, overcoming inertia
requires a strong connection to the group identity, or against an individual,
event, or political injustice [@Goldstone_1980; Goldstone_2011]. Political
resistance offers participants the opportunity to “lose themselves” and
relinquish their individual identity in favor of identification with a larger
group identity, thus contributing to a cohesive group ready and mobilized for
action [@Hoffer_2011. Previous research on social movements supports the idea
that participants are likely to join movements when they know and have a
personal connection to someone else already involved [@McAdam_1986. The type of
collective contentious behavior that manifests in society is a function of the
available political opportunity structures. The political opportunity structure,
as defined by Kitschelt (1986) are “comprised of specific configuration of
resources, institutional arrangements and historical precedents for social
mobilization, which facilitate the development of protest movements in some
instances and constrain them in others [@Kitschelt_1986].” What emerges as
strategies and tactics in contentious politics is in essence, a function of what
the system makes possible. This includes the number of people and their
disposition, i.e. nonviolent or violent, as well as their internal structure.
The more cohesive a group is, the greater threat they pose to the establishment.

>>>>

### Lexical and Interlocutor Cohesion

@@

Given the importance of lexical and interlocutor cohesion for messaging in
political movements, lexical cohesion may be an important bit of trace data for
researchers interested in leveraging social media data to analyze social
movements. Similar to large-scale studies of sociocultural movements [sometimes
referred to as *culturomics*; e.g., @michel2011quantitative], political
scientists can track the ways in which conversations on social media coalesce or
diffuse around similar content words. In this way, it matters relatively little
whether the interlocutors have precisely the same views: The point is to
understand whether the society-level conversation is converging on similar
language.

### Social and Political Cohesion

@@

At the same time, behavioral similarity in certain contexts has been
connected to important psychological and social effects. For example, shared
language is associated with rapport [both in face-to-face and computer-mediated
conversations; @niederhoffer2002linguistic; @riordan2011evidence] and
can even be connected to better performance on shared tasks [as long as
the language is relevant to the task at hand; @fusaroli2012coming]. Shared
movement is associated with greater group bonding and greater willingness
to contribute to the group [@wiltermuth2009synchrony], and shared identity 
is associated with a greater willingness to adopt the group's goals and
motivations [@walton2012mere].

We should expect to see increased linguistic cohesion among people participating
in online discussions about an idea, event, individual, or movement. Given that
Syrian protesters are known to have coordinated their activity through Twitter
by discussing their own events (e.g., planned protests, active demonstrations)
and their reactions to events happening around them (e.g., government reactions,
international action or inaction), the cohesion in social media should be
expected to correlate with the social cohesion of the group [@zeitzoff2011using].

This social cohesion can have real and wide-ranging impacts on a political
movement and the people in it. A socially cohesive group is inherently more
threatening to the existing regime because it reflects collective opposition
momentum and the professionalization of potential alternate leadership. Groups
with high social cohesion are more difficult to divide and conquer, and more
cohesive messages may also garner increased international support. Global
insurgency movements exhibit similar patterns in their social structure
[@bohorquez2009common]. Each violent group has a stable state. Exogenous events
can create perturbations that push the group to become either more cohesive or
more fragmented. Rather than frenetic, patchwork messages via Twitter, emergent
or intentional cohesive online messaging---using similar language, paralleling
in-person messaging [@Windsor2020]---are a sign of increased organizational
structure, inciting a government response.

Thus, lexical and interpersonal cohesion are hypothesized to be tied to social
and political cohesion. Shared language in an online environment may lead to the
development or expression of a shared identity [@wiltermuth2009synchrony]. This,
in turn, may drive individuals’ propensity to engage in similar coordinated
behavior on behalf of the group [@wiltermuth2009synchrony; @walton2012mere].
Taken together, these two processes may grow movement participants' weak and
strong ties [@granovetter1977strength; @gladwell2010revolution;
@mcadam1986recruitment], strengthening the movement overall by expanding and
deepening heterogenous ties. Therefore, 
in the context of social conflict, the linguistic cohesion
of the online conversation should correspond to the degree of
social cohesion on the ground.

Steinert-Threlkeld [-@steinert2017spontaneous] modeled a similar relationship
between social media and social mobilization alongside event data. This
groundbreaking work demonstrates that participants on the fringe of a social
movement contribute substantially to the overall level of protest. In addition
to providing new methods to this growing field of study, we build on this
previous work by demonstrating how sociolinguistic cohesion evolves in
response to negative and positive exogenous events.

Political and social cohesion in the world can be modeled using event data which
encodes who does what to whom on a given day. This could include events such as
the government imposing a blockade or fighting with artillery or tanks on
a civilian group. CAMEO (Conflict and Mediation Event Observations) encode
a set of activities with numerical values, rescaled to reflect a
conflict-cooperation scale [@schrodt2006twenty; @goldstein1992conflict]. These event
codes include a range of behaviors, from making public statements and appeals,
to demanding, threatening, protesting, and using unconventional mass violence.

#### Cohesion in Social Movements and Civil Wars
People join social and political resistance movements because they feel a sense
of connection to the cause, and because they have been sufficiently convinced to
participate. Given that people are generally risk averse, overcoming inertia
requires a strong connection to the group identity, or against an individual,
event, or political injustice [@Goldstone_1980; Goldstone_2011]. Political
resistance offers participants the opportunity to “lose themselves” and
relinquish their individual identity in favor of identification with a larger
group identity, thus contributing to a cohesive group ready and mobilized for
action [@Hoffer_2011. Previous research on social movements supports the idea
that participants are likely to join movements when they know and have a
personal connection to someone else already involved [@McAdam_1986. The type of
collective contentious behavior that manifests in society is a function of the
available political opportunity structures. The political opportunity structure,
as defined by Kitschelt (1986) are “comprised of specific configuration of
resources, institutional arrangements and historical precedents for social
mobilization, which facilitate the development of protest movements in some
instances and constrain them in others [@Kitschelt_1986].” What emerges as
strategies and tactics in contentious politics is in essence, a function of what
the system makes possible. This includes the number of people and their
disposition, i.e. nonviolent or violent, as well as their internal structure.
The more cohesive a group is, the greater threat they pose to the establishment.
